### 2017-08-31 paper.tex
:books: Draft

\Abstract{This paper investigates the effect of permeability on the runout characteristics of collapse of  granular columns on slopes in fluid. Two-dimensional sub-grain scale numerical simulations are performed to understand the flow dynamics of granular collapse in fluid. The Discrete Element (DEM) technique is coupled with the Lattice Boltzmann Method (LBM), for fluid-grain interactions, to understand the evolution of submerged granular flows. The fluid phase is simulated using Multiple-Relaxation-Time LBM (LBM-MRT) for numerical stability. In order to simulate interconnected pore space in 2D, a reduction in the radius of the grains (hydrodynamic radius) is assumed during LBM computations. The collapse of granular column in fluid is compared with the dry cases to understand the effect of fluid on the runout behaviour. A parametric analysis is performed to assess the influence of the granular characteristics (initial packing) on the evolution of flow and run-out distances for slope angles of 5°. In order to understand the effect of permeability on granular flow down a slope angle of 5°, the collapse of a granular column with an initial aspect ratio of 0.8 is simulated with different permeabilities. The hydrodynamic radius of a loosely packed granular column is varied from r = 0.7 R (high permeability), 0.75 R, 0.8 R, 0.85 R to 0.9 R (low permeability). The granular flow dynamics is investigated by analysing the effect of hydroplaning, water entrainment and viscous drag on the granular mass. The mechanism of energy dissipation, shape of the flow front, water entrainment and evolution of packing density is used to explain the difference in the flow characteristics of granular column collapse in fluid.}

@@

>>>>

\Abstract{This paper investigates the effect of permeability on the runout characteristics of collapse of  granular columns on slopes in fluid. Two-dimensional sub-grain scale numerical simulations are performed to understand the flow dynamics of granular collapse in fluid. The Discrete Element (DEM) technique is coupled with the Lattice Boltzmann Method (LBM), for fluid-grain interactions, to understand the evolution of submerged granular flows. The fluid phase is simulated using Multiple-Relaxation-Time LBM (LBM-MRT) for numerical stability. In order to simulate interconnected pore space in 2D, a reduction in the radius of the grains (hydrodynamic radius) is assumed during LBM computations. The collapse of granular column in fluid is compared with the dry cases to understand the effect of fluid on the runout behaviour. A parametric analysis is performed to assess the influence of the granular characteristics (initial packing) on the evolution of flow and run-out distances for slope angle of \SI{5}{\degree}. In order to understand the effect of permeability on granular flow down a slope angle of \SI{5}{\degree}, the collapse of a granular column with an initial aspect ratio of 0.8 is simulated with different permeabilities. The hydrodynamic radius of a loosely packed granular column is varied from r = 0.7 R (high permeability), 0.75 R, 0.8 R, 0.85 R to 0.9 R (low permeability). The granular flow dynamics is investigated by analysing the effect of hydroplaning, water entrainment and viscous drag on the granular mass. The mechanism of energy dissipation, shape of the flow front, water entrainment and evolution of packing density is used to explain the difference in the flow characteristics of granular column collapse in fluid.}

@@

\section{Summary and Conclusions}

Two-dimensional LB-DEM simulations were performed to understand the effect of initial volume fraction on the behaviour of granular column collapse on slopes submerged in fluid. The cases were run with a granular material with relatively low permeability by adopting varying hydraulic radius. By doing so, it was possible to examine the effect of coupled interaction of pore pressure and interparticle force during different stages of granular column collapse.

LBM-DEM simulation results show that a dense granular column behavior was initially affected by the development of negative pore pressure, which resulted in increased interparticle shear resistance. This caused the delay in initiating the mass flow and reduced the kinetic energy to move laterally. During the runout phase, the flow front had an acute angle creating more vortices and drag forces along the top of the granular mass, which resisted the movement of the granular mass. This in turn resulted in shorter runout distances when compared to the dry granular column collapse cases.

In the loose granular column collapse cases, positive pore pressures were observed inside the granular body as it sheared during the initiation phase. This allowed the body to initiate the movement faster than the dense column cases. During the runout phase, water was entrained at the flow front due to its particular parabolic shape and the packing density decreased, causing reduction in the interparticle forces at the flow front and creating some lubrication-like effect. The friction between the moving body and the bottom surface was also reduced to zero as there was no effective stress acting on the sliding plane (i.e. hydroplaning). The combined effect of positive pore pressure development, water entrainment in the front body and the hydroplaning resulted in longer runout distance compared to the dry granular column collapse cases that had the same loose packing density. An opposite behaviour was observed in the dense granular column cases, which shows the importance of initial packing density in the development of runout mechanisms.

In both dense and loose granular column cases, the final packing density was the same, indicating some critical state like behaviour, in which the initial packing memory is lost after large shearing of granular material.

\section{Acknowledgements}
The authors would like to thank Professor Farhang Radjai, LMGC, Montpellier, France for stimulating discussions regarding this work. The first author would like to thank the Cambridge Commonwealth, Overseas Trust and the Shell-Cambridge-Brazil collaboration for the financial support to pursue this research.
